Rheologic constraints on the upper mantle from five years of postseismic deformation following the El Mayor-Cucapah earthquake

### 2016-04-22 JGRSubmission/ElMayor.tex
commit before push

\begin{figure}
\includegraphics[scale=0.8]{Figures/ContextMap} 
\caption{Map of the region considered in this study.  The large focal mechanism is for the El Mayor-Cucapah earthquake, and the three small focal mechanisms are for the Ocotillo earthquake and the two main shocks during the Brawley swarm.  The black dots indicate the locations of GPS stations used in this study.  The fault geometry used in this study is shown in cyan where dashed lines indicate buried edges of the fault segments.  The green and red boxes demarcate the extent of the near-field and far-field maps (Figures \ref{fig:NearField} and \ref{fig:FarField}).  Stations inside the blue sector, which highlights the area within 10$^\circ$ of the El Mayor-Cucapah P axis, are used in Figures \ref{fig:RecordSectionElastic}, \ref{fig:RecordSection1}, and \ref{fig:RecordSection2}}.       
\label{fig:ContextMap}
\end{figure}

\section{Data Processing}\label{sec:Data}

We use continuous GPS position time series provided by University Navstar Consortium (UNAVCO) for stations within a 400 km radius about the El Mayor-Cucapah epicenter. We collectively describe the coseismic and postseismic displacements resulting from the El Mayor-Cucapah earthquake as $u_\mathrm{post}(t)$.  We consider the GPS position time series $u_\mathrm{obs}(t)$ to be the combination of $u_\mathrm{post}(t)$, secular tectonic deformation, annual and semi-annual oscillations, and coseismic offsets from significant earthquakes over the time span of this study.  The June 14, 2010, Mw5.8 Ocotillo earthquake and the Brawley swarm, which included an Mw5.5 and an Mw5.4 event on August 26, 2012, (Figure \ref{fig:ContextMap}) are the only earthquakes that produced noticeable displacements in any of the time series.  We treat the displacements resulting from the Brawley swarm as a single event because the daily solutions provided by UNAVCO cannot resolve the separate events.  Although the Ocotillo earthquake had its own series of aftershocks \citep{Hauksson2011}, neither the Ocotillo earthquake nor the Brawley swarm produced detectable postseismic deformation.  We model displacements resulting from these events with only a Heaviside function, $H(t)$, describing the coseismic offsets.  We then model $u_\mathrm{obs}(t)$ as 
\begin{equation}
  u_\mathrm{obs}(t) = u_\mathrm{pred}(t) + \epsilon,
\end{equation}
where
\begin{equation}\label{TimeSeriesModel}
  \begin{split}  
    u_\mathrm{pred}(t) = &u_\mathrm{post}(t)H(t-t_\mathrm{emc}) + c_0 + c_1t + \\
                         &c_2\sin(2\pi t) + c_3\cos(2\pi t) + c_4\sin(4\pi t) + c_5\cos(4\pi t) + \\
                         &c_6H(t-t_\mathrm{oc}) + c_7H(t-t_\mathrm{bs}).
  \end{split}
\end{equation}
In the above equations, $t_\mathrm{emc}$, $t_\mathrm{oc}$ and $t_\mathrm{bs}$ are the times of the El Mayor-Cucapah earthquake, Ocotillo earthquake, and the Brawley swarm, respectively, $c_0$ through $c_7$ are unknown coefficients, and $\epsilon$ is the observation noise. We only estimate jumps associated with the Ocotillo earthquake and Brawley swarm for stations within 40 km of their epicenters.

@@

Studies of postseismic deformation typically assume a parametric form for $u_\mathrm{post}(t)$, such as one with a logarithmic or exponential time dependence \citep[e.g.][]{Savage2005a}.  However, by assuming a logarithmic or exponential form of $u_\mathrm{post}(t)$ we run the risk of over fitting the GPS time series and inferring a non-existent postseismic signal. We therefore do not assume any parametric form for $u_\mathrm{post}(t)$ and rather treat it as integrated Brownian motion, so that 
\begin{equation}
    \dot{u}_\mathrm{post}(t) = \sigma^2\int_0^t w(s) ds,
\end{equation}    
where $w(t)$ is white noise and the variance of $\dot{u}_\mathrm{post}(t)$ increases linearly with time by a factor of $\sigma^2$. We use a Kalman filtering approach to estimate $u_\mathrm{post}(t)$ and the unknown parameters in eq. (\ref{TimeSeriesModel}).  In the context of Kalman filtering, our time varying state vector is
\begin{equation}
    \mathbf{X}(t) = [u_\mathrm{post}(t),\dot u_\mathrm{post}(t), c_0, ..., c_7]
\end{equation}
and eq. (\ref{TimeSeriesModel}) is the observation function which maps the state vector to the GPS observations. We initiate the Kalman filter by assuming a prior estimate of $\mathbf{X}(t)$ at the first time epoch, denoted $\mathbf{X}_{1|0}$, which has a sufficiently large covariance, denoted $\mathbf{\Sigma}_{1|0}$, to effectively make our prior uninformed.  For each time epoch $t_i$, Bayesian linear regression is used to incorporate GPS derived estimates of displacement with our prior estimate of the state $\mathbf{X}_{i|i-1}$ to form a posterior estimate of the state $\mathbf{X}_{i|i}$, which has covariance $\mathbf{\Sigma}_{i|i}$.  We then use the posterior estimate of the state at time $t_i$ to form a prior estimate of the state at time $t_{i+1}$ through the transition function
\begin{equation}\label{predict}
  \mathbf{X}_{i+1|i} = \mathbf{F}_{i+1}\mathbf{X}_{i|i} + \mathbf{\delta}_{i+1}, 
\end{equation}
where 
\begin{equation}
  \mathbf{F}_{i+1} = 
  \left[
  \begin{array}{ccc}
    1           & (t_{i+1} - t_i) & \mathbf{0}\\
    0           & 1              & \mathbf{0}\\
    \mathbf{0}  & \mathbf{0}     & \mathbf{I}
  \end{array}
  \right]
\end{equation}
and $\mathbf{\delta}_{i+1}$ is the process noise, which has zero mean and covariance described by
\begin{equation}
  \mathbf{Q}_{i+1} = 
  \sigma^2 \left[
  \begin{array}{ccc}
  \frac{(t_{i+1} - t_i)^3}{3} & \frac{(t_{i+1} - t_{i})^2}{2} & \mathbf{0}\\
  \frac{(t_{i+1} - t_i)^2}{2} & (t_{i+1} - t_{i}) & \mathbf{0}\\ 
  \mathbf{0} & \mathbf{0} & \mathbf{0}
  \end{array}
  \right].
\end{equation}
The covariance of the new prior state, $\mathbf{X}_{i+1|i}$, is then described by
\begin{equation}
  \mathbf{\Sigma}_{i+1|i} = \mathbf{F}_{i+1}\mathbf{\Sigma}_{i|i}\mathbf{F}^T_{i+1} + \mathbf{Q}_{i+1}.
\end{equation}
This process is repeated for each of the $N$ time epochs at which point we use Rauch-Tung-Striebel smoothing \citep{Rauch1965} to find $\mathbf{X}_{i|N}$, which is an estimate of the state at time $t_i$ that incorporates GPS observation for all $N$ time epochs.  Our final estimates of $u_\mathrm{post}(t)$ are used in subsequent analysis, while the remaining components of the state vector are considered nuisance parameters. In the interests of computational tractability, we down sample our smoothed time series from daily solutions down to weekly solutions.

@@

\subsection{Elastic Postseismic Inversion}\label{sec:ElasticInversion}    
We consider a variety of rheologic models for the lower crust and upper mantle. The simplest rheologic model is to consider them to be effectively elastic and isotropic.  In such case, the rheologic parameters consist of the Lam\'e parameters, $\lambda$ and $\mu$, which are reasonably well known and we use the same values used by \citet{Wei2011} throughout this paper (Table \ref{tab:MaterialProperties}).  The only unknown is the distribution of fault slip, which can be easily estimated from postseismic deformation through linear least squares.  Using a subset of the GPS stations considered in this study, \citet{Rollins2015} found that three years of postseismic deformation following the El Mayor-Cucapah earthquake can be explained with afterslip on the coseismic fault plane without requiring any viscoelastic relaxation. We also perform an elastic slip inversion, but we use GPS stations within a larger radius about the El Mayor-Cucapah epicenter (400 km instead of ${\sim}200$ km). Our forward problem describing predicted postseismic deformation $u_\mathrm{pred}$ in terms of time dependent fault slip, $s$, is
\begin{equation}\label{eq:ElasticForward}
  u_\mathrm{pred}(x,t) = \int_F s(\xi,t)g(x,\xi)d\xi, 
\end{equation}           
where $F$ denotes the fault and $g(x,\xi)$ is the elastic Green's function describing displacement at surface position $x$ resulting from slip at $\xi$ on the fault.  We estimate coseismic slip and the rate of afterslip at the discrete time intervals, 0.0-0.125, 0.125-0.25, 0.5-1.0, 1.0-2.0, 2.0-3.0, 3.0-4.0, and 4.0-5.0 years after the earthquake.  Each fault segment is discretized into roughly 4 km by 4 km patches and we estimate a strike-slip and thrust component of slip for each patch. We impose that the direction of slip and slip rate are within 45$^\circ$ of right-lateral. We also add zeroth-order Tikhonov regularization so that our solution for $s$ satisfies
\begin{equation}\label{eq:ElasticObjective}
  \min_s \left(\left|\left|\frac{u_\mathrm{pred}(s) - u_\mathrm{post}}                
                                {\sigma_\mathrm{post}}\right|\right|_2^2 + 
                                \lambda_s||s||_2^2\right),
\end{equation}
where $\sigma_\mathrm{post}$ is the uncertainty on postseismic displacements and $\lambda_s$ is a penalty parameter which is chosen with a trade-off curve.  We use Pylith \citep{Aagaard2013} to compute the Green's functions for this inversion as well as for the remaining inversions in this paper.

@@

For any linear viscoelastic rheology of the crust and mantle, postseismic displacements resulting from time dependent fault slip can be described as  
\begin{equation}\label{GeneralForward}
  u_\mathrm{pred}(x,t) = \int_F s(\xi,t)g(x,\xi)d\xi + 
           \int_0^t\int_F s(\xi,\tau) f(t-\tau,x,\xi) d\xi d\tau,
\end{equation}
where $f(t,x,\xi)$ describes the time-dependent velocity at $x$ resulting from viscoelastic relaxation of stresses induced by slip at $\xi$. $f$ is a function of $\lambda$, $\mu$, and any additional rheologic parameters controlling the viscoelastic response, which are generally not well known. Schematic representations of the viscoelastic rheologic models considered in this study are shown in Figure \ref{fig:Rheology}.  We discuss these rheologic models and their use in geophysical studies in Section \ref{Discussion}.

@@

In order to greatly simplify the inverse problem, we use the method described in \citet{Hines2016} to constrain an initial effective viscosity structure from the early postseismic deformation.  Our method utilizes the fact that coseismic stresses throughout the crust and upper mantle depend on the instantaneous elastic parameters and are independent of the viscoelastic parameters which we wish to estimate. Immediately following an earthquake, each parcel will have a strain rate that is proportional to the coseismic stress and inversely proportional to the parcel's effective viscosity, 
$\eta_\mathrm{eff}$.  Using one-dimensional viscoelastic models, we define the effective viscosity as
\begin{equation}
  \eta_\mathrm{eff} = \left.\frac{\sigma}{\dot{\varepsilon}}\right|_{t=0},
\end{equation}
where $\sigma$ is an applied stress at $t=0$ and $\dot\varepsilon$ is the resulting strain rate.  Figure \ref{fig:Rheology} shows how $\eta_\mathrm{eff}$ relates to the parameters for various linear viscoelastic rheologies.   We can deduce that the initial rate of surface deformation resulting from viscoelastic relaxation is a summation of the surface deformation resulting from relaxation in each parcel, scaled by the reciprocal of the parcel's effective viscosity.  That is to say   
\begin{equation}\label{InitialRate}
  f(0,x,\xi) = \int_L \frac{h(x,\xi,\zeta)}{\eta_\mathrm{eff}(\zeta)} d\zeta, 
\end{equation}
where $L$ denotes the crust and mantle and $h(x,\xi,\zeta)$ describes the initial rate of deformation resulting from viscoelastic relaxation at $\zeta$ induced by slip at $\xi$. We can combine eq. (\ref{InitialRate}) with eq. (\ref{GeneralForward}) to get a first order approximation for early postseismic deformation,
\begin{equation}\label{ApproxForward}
  u_\mathrm{pred}(x,t) \approx \int_F s(\xi,t)g(x,\xi)d\xi + 
           \int_0^t\int_F\int_L \frac{s(\tau,\xi)}{\eta_\mathrm{eff}(\zeta)} h(x,\xi,\zeta) d\zeta d\xi d\tau,
\end{equation}
which is valid for as long as the rate of deformation resulting from viscoelastic relaxation is approximately constant.  Although eq. (\ref{ApproxForward}) may only be valid for a short portion of the postseismic period, its utility becomes apparent when noting that $g$ and $h$ are only functions of the fault geometry and instantaneous elastic properties, $\lambda$ and $\mu$, and thus $g$ and $h$ can be computed numerically as a preprocessing step.  The forward problem in eq. (\ref{ApproxForward}) can then be rapidly evaluated for any realization of $s$ and $\eta_{\mathrm{eff}}$.  This is in contrast to evaluating the full forward problem, eq. (\ref{GeneralForward}), numerically for each realization of $s$ and the unknown rheologic properties.

@@

We estimate coseismic slip, afterslip, and effective viscosities by solving 
\begin{equation}\label{ObjectiveFunction}
 \mathrm{min}_{s,\eta_\mathrm{eff}} \left(\left|\left|
 \frac{u_\mathrm{pred}(s,\eta_\mathrm{eff}) - u_\mathrm{post}}
 {\mathbf{\sigma_\mathrm{post}}}\right|\right|_2^2 + 
 \lambda_s||s||_2^2 + 
 \lambda_\eta||\nabla \eta_{\mathrm{eff}}^{-1}||_2^2\right),
\end{equation} 
where $u_\mathrm{post}$ consists of the first 0.8 years of postseismic deformation and $u_\mathrm{pred}$ are the predicted displacements from eq. (\ref{ApproxForward}).  Due to inherent non-uniqueness, we have added zeroth-order Tikhonov regularization to estimates of $s$ and second-order Tikhonov regularization to estimates of effective fluidity $\eta_\mathrm{eff}^{-1}$. The degree to which we impose the regularization on slip and fluidity is controlled by the penalty parameters $\lambda_s$ and $\lambda_\eta$, which are chosen with trade-off curves.  Our goal here is to get a prior constraint on $\eta_{\mathrm{eff}}$ to minimize the amount of searching we have to do when describing the postseismic deformation over the full five years, which we do in Section \ref{sec:FullInversion}.  Estimates of $s$ made here will not be used in Section \ref{sec:FullInversion}, and so the motivation behind adding regularization to $s$ is to ensure that the slip driving viscoelastic relaxation in eq. (\ref{ApproxForward}) is sensible.

@@

In the previous section, we used the inverse method from \citet{Hines2016} to constrain the effective viscosity structure required to explain the first 0.8 years of postseismic deformation. In this section, we use these effective viscosities as a prior constraint when searching for models which are capable of describing the available five years of postseismic data, where our forward problem is now eq. (\ref{GeneralForward}) rather than the approximation given by eq. (\ref{ApproxForward}).  We perform a series of fault slip inversions assuming a variety of rheologies for the lower crust and upper mantle which are consistent with our findings from Section \ref{sec:InitialInversion}.  We appraise each model using the mean chi-squared value,
\begin{equation}\label{eq:Misfit}
  \bar\chi^2 = \frac{1}{N}\left|\left|\frac{u_\mathrm{pred} - u_\mathrm{obs}}{\sigma_\mathrm{obs}}\right|\right|_2^2,
\end{equation}
where $N$ is the number of observations.

@@

\section*{Acknowledgements}
We thank Andy Freed for an illuminating discussion on the data used in this study.  This material is based on EarthScope Plate Boundary Observatory data services provided by UNAVCO through the GAGE Facility with support from the National Science Foundation (NSF) and National Aeronautics and Space Administration (NASA) under NSF Cooperative Agreement No. EAR-1261833.  This material is based upon work supported by the National Science Foundation under Grant Numbers EAR 1045372 and EAR 1245263.

>>>>

%\begin{figure}
%\includegraphics[scale=0.8]{Figures/ContextMap} 
%\caption{Map of the region considered in this study.  The large focal mechanism is for the El Mayor-Cucapah earthquake, and the three small focal mechanisms are for the Ocotillo earthquake and the two main shocks during the Brawley swarm.  The black dots indicate the locations of GPS stations used in this study.  The fault geometry used in this study is shown in cyan where dashed lines indicate buried edges of the fault segments.  The green and red boxes demarcate the extent of the near-field and far-field maps (Figures \ref{fig:NearField} and \ref{fig:FarField}).  Stations inside the blue sector, which highlights the area within 10$^\circ$ of the El Mayor-Cucapah P axis, are used in Figures \ref{fig:RecordSectionElastic}, \ref{fig:RecordSection1}, and \ref{fig:RecordSection2}}.       
%\label{fig:ContextMap}
%\end{figure}

\section{Data Processing}\label{sec:Data}
We use continuous GPS position time series provided by University Navstar Consortium (UNAVCO) for stations within a 400 km radius about the El Mayor-Cucapah epicenter. We collectively describe the coseismic and postseismic displacements resulting from the El Mayor-Cucapah earthquake as $u_\mathrm{post}(t)$.  We consider the GPS position time series $u_\mathrm{obs}(t)$ to be the combination of $u_\mathrm{post}(t)$, secular tectonic deformation, annual and semi-annual oscillations, and coseismic offsets from significant earthquakes over the time span of this study.  The June 14, 2010, Mw5.8 Ocotillo earthquake and the Brawley swarm, which included an Mw5.5 and an Mw5.4 event on August 26, 2012, (Figure \ref{fig:ContextMap}) are the only earthquakes that produced noticeable displacements in any of the time series.  We treat the displacements resulting from the Brawley swarm as a single event because the daily solutions provided by UNAVCO cannot resolve the separate events.  Although the Ocotillo earthquake had its own series of aftershocks \citep{Hauksson2011}, neither the Ocotillo earthquake nor the Brawley swarm produced detectable postseismic deformation.  We model displacements resulting from these events with only a Heaviside function, $H(t)$, describing the coseismic offsets.  We then model $u_\mathrm{obs}(t)$ as

\begin{equation}
  u_\mathrm{obs}(t) = u_\mathrm{pred}(t) + \epsilon,
\end{equation}
where

\begin{equation}\label{TimeSeriesModel}
  \begin{split}  
    u_\mathrm{pred}(t) = &u_\mathrm{post}(t)H(t-t_\mathrm{emc}) + c_0 + c_1t + \\
                         &c_2\sin(2\pi t) + c_3\cos(2\pi t) + c_4\sin(4\pi t) + c_5\cos(4\pi t) + \\
                         &c_6H(t-t_\mathrm{oc}) + c_7H(t-t_\mathrm{bs}).
  \end{split}
\end{equation}
In the above equations, $t_\mathrm{emc}$, $t_\mathrm{oc}$ and $t_\mathrm{bs}$ are the times of the El Mayor-Cucapah earthquake, Ocotillo earthquake, and the Brawley swarm, respectively, $c_0$ through $c_7$ are unknown coefficients, and $\epsilon$ is the observation noise. We only estimate jumps associated with the Ocotillo earthquake and Brawley swarm for stations within 40 km of their epicenters.

@@

Studies of postseismic deformation typically assume a parametric form for $u_\mathrm{post}(t)$, such as one with a logarithmic or exponential time dependence \citep[e.g.][]{Savage2005a}.  However, by assuming a logarithmic or exponential form of $u_\mathrm{post}(t)$ we run the risk of over fitting the GPS time series and inferring a non-existent postseismic signal. We therefore do not assume any parametric form for $u_\mathrm{post}(t)$ and rather treat it as integrated Brownian motion, so that

\begin{equation}
    \dot{u}_\mathrm{post}(t) = \sigma^2\int_0^t w(s) ds,
\end{equation}    
where $w(t)$ is white noise and the variance of $\dot{u}_\mathrm{post}(t)$ increases linearly with time by a factor of $\sigma^2$. We use a Kalman filtering approach to estimate $u_\mathrm{post}(t)$ and the unknown parameters in eq. (\ref{TimeSeriesModel}).  In the context of Kalman filtering, our time varying state vector is

\begin{equation}
    \mathbf{X}(t) = [u_\mathrm{post}(t),\dot u_\mathrm{post}(t), c_0, ..., c_7]
\end{equation}
and eq. (\ref{TimeSeriesModel}) is the observation function which maps the state vector to the GPS observations. We initiate the Kalman filter by assuming a prior estimate of $\mathbf{X}(t)$ at the first time epoch, denoted $\mathbf{X}_{1|0}$, which has a sufficiently large covariance, denoted $\mathbf{\Sigma}_{1|0}$, to effectively make our prior uninformed.  For each time epoch $t_i$, Bayesian linear regression is used to incorporate GPS derived estimates of displacement with our prior estimate of the state $\mathbf{X}_{i|i-1}$ to form a posterior estimate of the state $\mathbf{X}_{i|i}$, which has covariance $\mathbf{\Sigma}_{i|i}$.  We then use the posterior estimate of the state at time $t_i$ to form a prior estimate of the state at time $t_{i+1}$ through the transition function

\begin{equation}\label{predict}
  \mathbf{X}_{i+1|i} = \mathbf{F}_{i+1}\mathbf{X}_{i|i} + \mathbf{\delta}_{i+1}, 
\end{equation}
where

\begin{equation}
  \mathbf{F}_{i+1} = 
  \left[
  \begin{array}{ccc}
    1           & (t_{i+1} - t_i) & \mathbf{0}\\
    0           & 1              & \mathbf{0}\\
    \mathbf{0}  & \mathbf{0}     & \mathbf{I}
  \end{array}
  \right]
\end{equation}
and $\mathbf{\delta}_{i+1}$ is the process noise, which has zero mean and covariance described by

\begin{equation}
  \mathbf{Q}_{i+1} = 
  \sigma^2 \left[
  \begin{array}{ccc}
  \frac{(t_{i+1} - t_i)^3}{3} & \frac{(t_{i+1} - t_{i})^2}{2} & \mathbf{0}\\
  \frac{(t_{i+1} - t_i)^2}{2} & (t_{i+1} - t_{i}) & \mathbf{0}\\ 
  \mathbf{0} & \mathbf{0} & \mathbf{0}
  \end{array}
  \right].
\end{equation}
The covariance of the new prior state, $\mathbf{X}_{i+1|i}$, is then described by

\begin{equation}
  \mathbf{\Sigma}_{i+1|i} = \mathbf{F}_{i+1}\mathbf{\Sigma}_{i|i}\mathbf{F}^T_{i+1} + \mathbf{Q}_{i+1}.
\end{equation}
This process is repeated for each of the $N$ time epochs at which point we use Rauch-Tung-Striebel smoothing \citep{Rauch1965} to find $\mathbf{X}_{i|N}$, which is an estimate of the state at time $t_i$ that incorporates GPS observation for all $N$ time epochs.  Our final estimates of $u_\mathrm{post}(t)$ are used in subsequent analysis, while the remaining components of the state vector are considered nuisance parameters. In the interests of computational tractability, we down sample our smoothed time series from daily solutions down to weekly solutions.

@@

\subsection{Elastic Postseismic Inversion}\label{sec:ElasticInversion}    
We consider a variety of rheologic models for the lower crust and upper mantle. The simplest rheologic model is to consider them to be effectively elastic and isotropic.  In such case, the rheologic parameters consist of the Lam\'e parameters, $\lambda$ and $\mu$, which are reasonably well known and we use the same values used by \citet{Wei2011} throughout this paper (Table \ref{tab:MaterialProperties}).  The only unknown is the distribution of fault slip, which can be easily estimated from postseismic deformation through linear least squares.  Using a subset of the GPS stations considered in this study, \citet{Rollins2015} found that three years of postseismic deformation following the El Mayor-Cucapah earthquake can be explained with afterslip on the coseismic fault plane without requiring any viscoelastic relaxation. We also perform an elastic slip inversion, but we use GPS stations within a larger radius about the El Mayor-Cucapah epicenter (400 km instead of ${\sim}200$ km). Our forward problem describing predicted postseismic deformation $u_\mathrm{pred}$ in terms of time dependent fault slip, $s$, is

\begin{equation}\label{eq:ElasticForward}
  u_\mathrm{pred}(x,t) = \int_F s(\xi,t)g(x,\xi)d\xi, 
\end{equation}           
where $F$ denotes the fault and $g(x,\xi)$ is the elastic Green's function describing displacement at surface position $x$ resulting from slip at $\xi$ on the fault.  We estimate coseismic slip and the rate of afterslip at the discrete time intervals, 0.0-0.125, 0.125-0.25, 0.5-1.0, 1.0-2.0, 2.0-3.0, 3.0-4.0, and 4.0-5.0 years after the earthquake.  Each fault segment is discretized into roughly 4 km by 4 km patches and we estimate a strike-slip and thrust component of slip for each patch. We impose that the direction of slip and slip rate are within 45$^\circ$ of right-lateral. We also add zeroth-order Tikhonov regularization so that our solution for $s$ satisfies

\begin{equation}\label{eq:ElasticObjective}
  \min_s \left(\left|\left|\frac{u_\mathrm{pred}(s) - u_\mathrm{post}}                
                                {\sigma_\mathrm{post}}\right|\right|_2^2 + 
                                \lambda_s||s||_2^2\right),
\end{equation}
where $\sigma_\mathrm{post}$ is the uncertainty on postseismic displacements and $\lambda_s$ is a penalty parameter which is chosen with a trade-off curve.  We use Pylith \citep{Aagaard2013} to compute the Green's functions for this inversion as well as for the remaining inversions in this paper.

@@

For any linear viscoelastic rheology of the crust and mantle, postseismic displacements resulting from time dependent fault slip can be described as

\begin{equation}\label{GeneralForward}
  u_\mathrm{pred}(x,t) = \int_F s(\xi,t)g(x,\xi)d\xi + 
           \int_0^t\int_F s(\xi,\tau) f(t-\tau,x,\xi) d\xi d\tau,
\end{equation}
where $f(t,x,\xi)$ describes the time-dependent velocity at $x$ resulting from viscoelastic relaxation of stresses induced by slip at $\xi$. $f$ is a function of $\lambda$, $\mu$, and any additional rheologic parameters controlling the viscoelastic response, which are generally not well known. Schematic representations of the viscoelastic rheologic models considered in this study are shown in Figure \ref{fig:Rheology}.  We discuss these rheologic models and their use in geophysical studies in Section \ref{Discussion}.

@@

In order to greatly simplify the inverse problem, we use the method described in \citet{Hines2016} to constrain an initial effective viscosity structure from the early postseismic deformation.  Our method utilizes the fact that coseismic stresses throughout the crust and upper mantle depend on the instantaneous elastic parameters and are independent of the viscoelastic parameters which we wish to estimate. Immediately following an earthquake, each parcel will have a strain rate that is proportional to the coseismic stress and inversely proportional to the parcel's effective viscosity, 
$\eta_\mathrm{eff}$.  Using one-dimensional viscoelastic models, we define the effective viscosity as

\begin{equation}
  \eta_\mathrm{eff} = \left.\frac{\sigma}{\dot{\varepsilon}}\right|_{t=0},
\end{equation}
where $\sigma$ is an applied stress at $t=0$ and $\dot\varepsilon$ is the resulting strain rate.  Figure \ref{fig:Rheology} shows how $\eta_\mathrm{eff}$ relates to the parameters for various linear viscoelastic rheologies.   We can deduce that the initial rate of surface deformation resulting from viscoelastic relaxation is a summation of the surface deformation resulting from relaxation in each parcel, scaled by the reciprocal of the parcel's effective viscosity.  That is to say

\begin{equation}\label{InitialRate}
  f(0,x,\xi) = \int_L \frac{h(x,\xi,\zeta)}{\eta_\mathrm{eff}(\zeta)} d\zeta, 
\end{equation}
where $L$ denotes the crust and mantle and $h(x,\xi,\zeta)$ describes the initial rate of deformation resulting from viscoelastic relaxation at $\zeta$ induced by slip at $\xi$. We can combine eq. (\ref{InitialRate}) with eq. (\ref{GeneralForward}) to get a first order approximation for early postseismic deformation,

\begin{equation}\label{ApproxForward}
  u_\mathrm{pred}(x,t) \approx \int_F s(\xi,t)g(x,\xi)d\xi + 
           \int_0^t\int_F\int_L \frac{s(\tau,\xi)}{\eta_\mathrm{eff}(\zeta)} h(x,\xi,\zeta) d\zeta d\xi d\tau,
\end{equation}
which is valid for as long as the rate of deformation resulting from viscoelastic relaxation is approximately constant.  Although eq. (\ref{ApproxForward}) may only be valid for a short portion of the postseismic period, its utility becomes apparent when noting that $g$ and $h$ are only functions of the fault geometry and instantaneous elastic properties, $\lambda$ and $\mu$, and thus $g$ and $h$ can be computed numerically as a preprocessing step.  The forward problem in eq. (\ref{ApproxForward}) can then be rapidly evaluated for any realization of $s$ and $\eta_{\mathrm{eff}}$.  This is in contrast to evaluating the full forward problem, eq. (\ref{GeneralForward}), numerically for each realization of $s$ and the unknown rheologic properties.

@@

We estimate coseismic slip, afterslip, and effective viscosities by solving

\begin{equation}\label{ObjectiveFunction}
 \mathrm{min}_{s,\eta_\mathrm{eff}} \left(\left|\left|
 \frac{u_\mathrm{pred}(s,\eta_\mathrm{eff}) - u_\mathrm{post}}
 {\mathbf{\sigma_\mathrm{post}}}\right|\right|_2^2 + 
 \lambda_s||s||_2^2 + 
 \lambda_\eta||\nabla \eta_{\mathrm{eff}}^{-1}||_2^2\right),
\end{equation} 
where $u_\mathrm{post}$ consists of the first 0.8 years of postseismic deformation and $u_\mathrm{pred}$ are the predicted displacements from eq. (\ref{ApproxForward}).  Due to inherent non-uniqueness, we have added zeroth-order Tikhonov regularization to estimates of $s$ and second-order Tikhonov regularization to estimates of effective fluidity $\eta_\mathrm{eff}^{-1}$. The degree to which we impose the regularization on slip and fluidity is controlled by the penalty parameters $\lambda_s$ and $\lambda_\eta$, which are chosen with trade-off curves.  Our goal here is to get a prior constraint on $\eta_{\mathrm{eff}}$ to minimize the amount of searching we have to do when describing the postseismic deformation over the full five years, which we do in Section \ref{sec:FullInversion}.  Estimates of $s$ made here will not be used in Section \ref{sec:FullInversion}, and so the motivation behind adding regularization to $s$ is to ensure that the slip driving viscoelastic relaxation in eq. (\ref{ApproxForward}) is sensible.

@@

In the previous section, we used the inverse method from \citet{Hines2016} to constrain the effective viscosity structure required to explain the first 0.8 years of postseismic deformation. In this section, we use these effective viscosities as a prior constraint when searching for models which are capable of describing the available five years of postseismic data, where our forward problem is now eq. (\ref{GeneralForward}) rather than the approximation given by eq. (\ref{ApproxForward}).  We perform a series of fault slip inversions assuming a variety of rheologies for the lower crust and upper mantle which are consistent with our findings from Section \ref{sec:InitialInversion}.  We appraise each model using the mean chi-squared value,

\begin{equation}\label{eq:Misfit}
  \bar\chi^2 = \frac{1}{N}\left|\left|\frac{u_\mathrm{pred} - u_\mathrm{obs}}{\sigma_\mathrm{obs}}\right|\right|_2^2,
\end{equation}
where $N$ is the number of observations.

@@

\section*{Acknowledgements}
We thank Andy Freed for an illuminating discussion on the data used in this study.  This material is based on EarthScope Plate Boundary Observatory data services provided by UNAVCO through the GAGE Facility with support from the National Science Foundation (NSF) and National Aeronautics and Space Administration (NASA) under NSF Cooperative Agreement No. EAR-1261833.  The data used in this study can be found at www.unavco.org. This material is based upon work supported by the National Science Foundation under Grant Numbers EAR 1045372 and EAR 1245263.
